Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-09-09 main.tex
completing edges

\begin{theorem}\label{theorem:starreflection}
    Let $\calT$ be an $n$-dimensional triangulation of an $n$-dimensional domain. 
    Let $S \in \calT$ be a simplex of dimension $k < n$,
    let $T \in \patch_{\calT}(S)$ be of dimension $n$,
    and let 
    \begin{align*}
        U := |\patch_{\calT}(S)| \setminus T.
    \end{align*}
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \Xi : T \rightarrow U
    \end{align*}
    satisfying
    \begin{gather*}
        \| \Jacobian\Xi \|          \leq \Cvier{n}(\calT),
        \quad 
        \| \Jacobian\Xi^{-1} \|     \leq \Cfive{n}(\calT),
        \\
        \| \det\Jacobian\Xi \|      \leq \Cdetvier{n}(\calT),
        \quad 
        \| \det\Jacobian\Xi^{-1} \| \leq \Cdetfive{n}(\calT),
    \end{gather*}
    where the constants on the right-hand sides are as stated in the proof. 
\end{theorem}
\begin{proof}
    We derive the estimate in several steps. 
    \begin{itemize}
        \item 
        Without loss of generality, the barycenter of $S$ is the origin. 
        % Let $I := U \cap T$, which is the union of $k$ different faces of $T$, with $1 \leq k \leq n$.
        %We let $J \subseteq \partial T$ be the union of the remaining $n+1-k$ faces of $T$.
        %So $J$ is within the boundary of the star $|\patch_{\calT}(S)|$ and is disjoint from $S$.
        We fix the subsimplex $S' \subseteq T$ that is complementary to $S$.
        Note that $J$ is the union of exactly those faces of $T$ that contain $S'$.

@@

        We introduce a new mapping $\Theta : |\calR| \rightarrow |\calR^c|$.
        Let $K \in \calR$ be an $n$-simplex and $K' \in \calR^{c}$ be constructed from $K$.
        We let $A_{K} : \Delta^{n} \rightarrow K$ and $A_{K'} : \Delta^{n} \rightarrow K'$
        be affine reference transformations
        that agree on the vertices common to $K$ and $K'$.
        We then define 
            \Theta_{|K} := A_{K'} \circ A_{K}^{-1}.
        Notice that $A_{K'}$ and $A_{K}$ have the same singular values. Thus
            \| \Jacobian \Theta_{|K}      \| &\leq \| A_{K'} \| \cdot \| A_{K}^{-1} \| \leq \kappa(A_{K}),
            \| \Jacobian \Theta_{|K}^{-1} \| &\leq \| A_{K'}^{-1} \| \cdot \| A_{K} \| \leq \kappa(A_{K}).
        \color{blue}
        By construction, 
            |\det( \Jacobian      \Theta_{|K} )|
            |\det( \Jacobian^{-1} \Theta_{|K} )|
        We $h_z \in \bbR^{n}$ be the height vector of $z_{S'}$ inside the simplex $K$.
        Clearly,
            \Theta_{|K}(x) 
            2 \hat z_{S'} \frac{\langle \hat h_z, x \rangle}{\langle \hat h_z, \hat z_{S'} \rangle}
        Indeed, we check that the right-hand side equals $x$ whenever $x$ lies in the plane orthgonal to $\hat h_z$,
        and when $x = z_{S'}$, then 
            -z_{S'} 
            z_{S'}
            2 \hat z_{S'} \frac{\langle \hat h_z, z_{S'} \rangle}{\langle \hat h_z, \hat z_{S'} \rangle}
        If we orthogonally decompose $z_{S'} = h_z + b_z$, then 
            \Theta_{|K}( h_z ) 
            h_z
            2 z_{S'} \frac{\langle h_z, h_z \rangle}{\langle h_z, z_{S'} \rangle}
            h_z
            2 z_{S'}
            - h_{z}
            2 b_{z}
        Let $\beta$ be the angle between $z_{S'}$ and $h_{z}$. 
        Then $\| h_z \| = \cos(\beta) \|z_{S'}\|$ and $\| b_z \| = \sin(\beta) \|z_{S'}\|$. 
        We study the singular values of the matrix 
            \begin{pmatrix}
            -1                 & 0
            -2 \|b_z\|/\|h_z\| & 1
            \end{pmatrix}
            \begin{pmatrix}
            -1             & 0
            -2 \tan(\beta) & 1
            \end{pmatrix}
            \begin{pmatrix}
            0             & \tan(\beta)
            -\tan(\beta) & 0
            \end{pmatrix}
            \begin{pmatrix}
            -1           & -\tan(\beta)
            \tan(\beta) & 1
            \end{pmatrix}
        It follows that $\det(\Theta_{|K}) = 1$. 
        Its singular values satisfy 
            ( \sigma_{\max} + \sigma_{\min} )/2
            \tan(\beta)
            \qquad 
            ( \sigma_{\max} - \sigma_{\min} )/2
            \sqrt{ 1 + \tan^{2}(\beta) }
        which means that 
            \sigma_{\max} = \sqrt{ 1 + \tan^{2}(\beta) } + \tan(\beta)
            \qquad 
            \sigma_{\min} = \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta)
        Let $\beta$ be the angle between $z_{S'}$ and $h_{z}$. 
        By the definition of the cosine, $\cos(\beta) \leq \| h_{z} \| / \| z_{S'} \|$. 
        Let $F_{z} \subseteq K$ be the face opposite to $z_{S'}$,
        which is also a face of $T$. Then $(n-k-1) h_{z}$ is the height vector of $F_{z}$ in $T$. 
        Thus 
            \frac{ \| h_{z} \| }{ \| z_{S'} \| }
            \frac{ (n-k-1)^{-1} \| (n-k-1) h_{z} \| }{ \| z_{S'} \| }
            \frac{ (n-k-1)^{-1} n \vol(T) }{ \| z_{S'} \| \vol(F_{z}) }
            \frac{n}{ n-k-1 }
            \frac{ \vol(T) }{ \diam(T)^{n} }
%             \\
%             \frac{ n \vol(K) }{ \| z_{S'} \| \vol(F_{z}) }
%             \frac{ \vol(T) }{ \| z_{S'} \| \vol(F_{z}) }
%             \frac{ \vol(T) }{ \diam(T)^{n} }

What are the singular values of a $2x2$ (lower) unit triangular matrix? Suppose we have the matrix
Suppose we have the matrix 
\begin{align*}
    \begin{pmatrix}
    1 & 0
    \\ 
    b & 1
    \end{pmatrix}
\end{align*}
Following up a generic formula found here, 
https://math.stackexchange.com/a/861774/3557
I decompose 
\begin{align*}
    \begin{pmatrix}
    1 & 0
    \\ 
    b & 1
    \end{pmatrix}
    =
    \frac 1 2 
    \begin{pmatrix}
    2 & -b
    \\ 
    b & 2
    \end{pmatrix}
    +
    \frac 1 2 
    \begin{pmatrix}
    0 & b
    \\ 
    b & 0
    \end{pmatrix}
\end{align*}
The singular values $w_1$ and $w_2$ are implicitly given by 
\begin{align*}
    \frac{ w_1 + w_2 }{2} = \frac 1 2 \sqrt{ 4 + b^2 }
    \\
    \frac{ w_1 - w_2 }{2} = \frac 1 2 |b|
\end{align*}
If we assume that the first value is at least the second value, 
then adding and subtracting these equations leads to 
\begin{align*}
    w_1 = \frac 1 2 \sqrt{ 4 + b^2 } + \frac 1 2 |b|
    \\
    w_2 = \frac 1 2 \sqrt{ 4 + b^2 } - \frac 1 2 |b|
\end{align*}
I would like to know whether that reasoning is correct

foo
\begin{align*}
    \lambda^2 = \frac{1}{2} \left( a + c \pm \sqrt{ a^2 - 2 a c + c^2 + 4b^2 } \right)
\end{align*}
\begin{align*}
    \lambda 
    &= 
    \frac{1}{\sqrt 2} \sqrt{ (1+t^2) + 1 \pm \sqrt{ (1+t^2)^2 - 2 (1+t^2)   +   1 + 4t^2 } } 
    \\
    &= 
    \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm \sqrt{ 1 + 2t^2 + t^4 - 2 - 2 t^2 + 1 + 4t^2 } } 
    \\
    &= 
    \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm \sqrt{          + t^4                 + 4t^2 } } 
    \\
    &= 
    \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm t \sqrt{ t^2 + 4 } } 
    .
\end{align*}
\begin{align*}
    \lambda = \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm t \sqrt{ t^2 + 4 } } 
\end{align*}
\begin{align*}
    \left( t^2 + 2 + t \sqrt{ t^2 + 4 } \right)
    \left( t^2 + 2 - t \sqrt{ t^2 + 4 } \right)
    =
    t^{4} + 4 t^{2} + 4
    -
    t^{2} ( t^{2} + 4 )
    =
    4
\end{align*}
\begin{align*}
    \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm t \sqrt{ t^2 + 4 } } 
    &=
    \sqrt{1+a^2/4} \pm a/2
    \\
    \frac 1 2 \left( t^2 + 2 \pm t \sqrt{ t^2 + 4 } \right)
    &=
    1+a^2/4
    \pm
    a \sqrt{1+a^2/4}
    + 
    a^2/4
    \\&=
    1
    +
    2a^2/4
    \pm
    a \sqrt{1+a^2/4}
    \\
    1 + \frac{t^2}{2} \pm t \sqrt{ \frac{t^2}{4} + 1 } 
    &=
    1
    +
    a^2/2
    \pm
    a \sqrt{1+a^2/4}
    .
\end{align*}

@@

        We identify the Lipschitz properties of this mapping by computing its Jacobian. 
        We write $\alpha_1, \alpha_2 \geq 0$ for the two angles between $h_1$ and $h_2$, respectively, and the vector $x$.
        In what follows, $y \in \bbR^{n}$.
        \begin{align*}
            \Jacobian F(x) \cdot y
            &=
            \frac{ \langle h_2, h_2 \rangle }{ \|h_2\| \|x\| \cos(\alpha_2) }
            \left( y - \frac{ \langle h_2, y \rangle }{ \langle h_2, x \rangle } x \right)
            \frac{ \langle x, h_1 \rangle }{ \langle h_1, h_1 \rangle }
            +
            x
            \otimes 
            \frac{ \| h_2 \|^{2} }{ \langle h_2, x \rangle }
            \frac{ h_1^{t} y }{ \| h_1 \|^{2} }
            \\&=
            \frac{ \langle \hat h_2, h_2 \rangle }{ \cos(\alpha_2) }
            \left( y - \frac{ \langle \hat h_2, y \rangle }{ \langle \hat h_2, \hat x \rangle } \hat x \right)
            \frac{ \langle \hat x, \hat h_1 \rangle }{ \langle h_1, \hat h_1 \rangle }
            +
            \hat x
            \otimes 
            \frac{ \| h_2 \| }{ \langle \hat h_2, \hat x \rangle }
            \frac{ \hat h_1^{t} y }{ \| h_1 \| }
            \\&=
            \frac{ \| h_2 \| }{ \cos(\alpha_2) }
            \left( y - \frac{ \langle \hat h_2, y \rangle }{ \cos(\alpha_2) } \hat x \right)
            \frac{ \cos(\alpha_1) }{ \| h_1 \| }
            +
            \hat x
            \otimes 
            \frac{ \| h_2 \| }{ \langle \hat h_2, \hat x \rangle }
            \frac{ \hat h_1^{t} y }{ \| h_1 \| }
            \\&=
            \frac{ \| h_2 \| }{ \| h_1 \| }
            \frac{ \cos(\alpha_1) }{ \cos(\alpha_2) }
            \left( 
                y 
                - 
                \frac{ \langle \hat h_2, y \rangle }{ \cos(\alpha_2) } \hat x 
                + 
                \frac{ \langle \hat h_1, y \rangle }{ \cos(\alpha_1) } \hat x 
            \right)
            .
        \end{align*}
        We define $a := \cos(\alpha_1)^{-1} \hat h_1 - \cos(\alpha_2)^{-1} \hat h_2$. 
        Notice that $a \perp \hat x$ by definition. The matrix $\Id + \hat x \otimes a^{t}$ is the identity over the orthogonal complement of the linear span of $a$ and $x$. Over that two-dimensional subspace, it can be represented by a triangular matrix.
        Its largest and the smallest singular values $\sigma_{max} \geq 1 \geq \sigma_{\min}$ are~\cite{blinn1996consider} 
        \begin{align*}
            \sigma_{\max} = \sqrt{ 1 + \frac{ \|a\|^{2} }{ 4 } } + \frac{ \|a\| }{2},
            \qquad 
            \sigma_{\min} = \sqrt{ 1 + \frac{ \|a\|^{2} }{ 4 } } - \frac{ \|a\| }{2}.     
        \end{align*}
        Notice that these are strictly monotonely increasing or decreasing, respectively, in $\|a\|$.
        We notice the obvious upper bound 
        \begin{align*}
            \|a\| 
            \leq 
            2 \max_{1 \leq i \leq 2} 
            \left( 
                \frac{ \hat h_i }{ \langle \hat x, \hat h_i \rangle } 
            \right)
            .
        \end{align*}

@@

        The diameters of the simplices in $\calK$ are already bounded by the maximum diameter in $\patch_{\calT}(S)$.
        The height $h_{K}$ is obtained from a height vector of a vertex in $\patch_{\calT}(S)$ by scaling with the factor $(k+1)^{-1}$.
        Every height of the origin in some simplex of $\calR'$ is obtained from a height vector of the origin in some simplex in $\calS'$
        by scaling with $(k+1)^{-1}$. Each of the latter height vectors has length bounded from below by the height of some vertex in $S$ within a simplex of $\calS$, multiplied by $\kappa(A)^{-2}$. The height of any vertex of $S$ in some simplex of $\calS$ is the same as the height $h_T$ of that vertex in the original simplex $T$. 
        In oder to estimate $\|x_{G}\|$, recall that $x = \lambda (-z_{S'}) + (1-\lambda) z_{x}$ for some $\lambda \in [0,1]$ and $z_{x} \in G$
        being in the face of $G$ that is opposite to $z_{x}$. The preimage of $x$ under $\Theta_{{G}}$ is $\lambda z_{S'} + (1-\lambda) z_{x}$. 
        It follows that $\|x_{G}\| \leq \diam(T)$.

        We know that whenever $T$ is a simplex with some face $F$ and $h_F$ is corresponding height vector, then 
        Recall that 
        \begin{align*}
            \frac{ \|h_{F}\|}{\diam(T)}
            = 
            \frac{ n \vol(T) }{ \vol(F) \diam(T) }
            \geq 
            \frac{ n \vol(T) }{ \diam(T)^{n} }
            \geq 
            (n-1)! \cdot \frac{ \diam(T)^{n} }{ n! \vol(T) }
            .
        \end{align*}
        We conclude that 
        \begin{gather*}
            \frac{ \diam(K) }{ \| h_{{G}} \| }
            \leq 
            (k+1)\kappa(A)^{2}
            \frac{ \diam(K) }{ \| h_{T} \| }
            \leq 
            (k+1)\kappa(A)^{2}
            \theta(\calT)
            \frac{ \diam(T) }{ \| h_{T} \| }
            \leq 
            (k+1)\kappa(A)^{2}
            (n-1)! \cdot \frac{ \diam(T)^{n} }{ n! \vol(T) }
            ,
            \\
            \frac{ \| x_{G} \| }{ \| h_{K} \| }
            \leq 
            (k+1)
            \frac{ \diam(T) }{ \| h_{K} \| }
            \leq 
            (k+1)
            \theta(\calT)
            \frac{ \diam(K) }{ \| h_{K} \| }
            \leq 
            (k+1)
            \theta(\calT)
            \sup_{ K \in \patch_{\calT}(S) }
            (n-1)! \cdot \frac{ \diam(K)^{n} }{ n! \vol(K) }
            .
        \end{gather*}
        \begin{gather*}
            \frac{ \diam(K) }{ \| h_{{K}} \| }
            \leq 
            (k+1)
            \sup_{ K \in \patch_{\calT}(S) }
            (n-1)! \cdot \frac{ \diam(K)^{n} }{ n! \vol(K) }
            ,
            \\
            \frac{ \| x_{G} \| }{ \| h_{G} \| }
            \leq 
            (k+1)\kappa(A)^{2}
            (n-1)! \cdot \frac{ \diam(T)^{n} }{ n! \vol(T) }
            .
        \end{gather*}

@@

        The vector $a$ is the difference between $\hat h_{G}$ and $\hat h_{K}$ after both are rescaled such that they lie on the affine hyperplane with normal vector $\hat x$. That difference has length at most $\tan(\alpha_{G}) + \tan(\alpha_{K})$.
        We calculate 
        \begin{align*}
            \|a\|
            = 
            \left\|
            \frac{\|x_{G}\|}{\|h_{G}\|} \hat h_{{G}} 
            - 
            \frac{\|x_{K}\|}{\|h_{K}\|} \hat h_K
            \right\|
            &
            \leq 
            \frac{\|x_{G}\|}{\|h_{G}\|} 
            + 
            \frac{\|x_{K}\|}{\|h_{K}\|} 
            \\&
            \leq 
            (k+1) 
            (n-1)! \cdot 
            \left( 1 + \kappa(A)^{2} \right)
            \sup_{ K \in \patch_{\calT}(S) }
            \frac{ \diam(K)^{n} }{ n! \vol(K) }
            .
        \end{align*}
        \todo{bound $\kappa(A)$ right away...}

@@

        We recall the upper bound $\sqrt{1+\theta^2} \leq 1 + \theta$ for any $\theta \geq 0$.

        \begin{align*}
            |\det(\Xi)|
            \leq 
            (k+1)^{n}
            (n-1)!^{n}
            \theta(\calT)^{n}
            \kappa(A)^{2n}
            \mu(\calT)
            ,
            \qquad 
            |\det(\Xi^{-1})|
            \leq 
            (k+1)^{n}
            (n-1)!^{n}
            \theta(\calT)^{n}
            \mu(\calT)
            .
        \end{align*}

        \color{black}

>>>>

\begin{theorem}\label{theorem:starreflection}
    Let $\calT$ be an $n$-dimensional triangulation of an $n$-dimensional domain. 
    Let $S \in \calT$ be a simplex of dimension $k < n$,
    let $T \in \patch_{\calT}(S)$ be of dimension $n$,
    and let 
    \begin{align*}
        U := |\patch_{\calT}(S)| \setminus T.
    \end{align*}
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \Xi : T \rightarrow U
    \end{align*}
    satisfying
    \begin{gather*}
        \| \Jacobian\Xi \|          \leq \Cvier{n}{k}(\calT),
        \quad 
        \| \Jacobian\Xi^{-1} \|     \leq \Cfive{n}{k}(\calT),
        \\
        \| \det\Jacobian\Xi \|      \leq \Cdetvier{n}{k}(\calT),
        \quad 
        \| \det\Jacobian\Xi^{-1} \| \leq \Cdetfive{n}{k}(\calT),
    \end{gather*}
    where the constants on the right-hand sides are as stated in the proof. 
\end{theorem}
\begin{proof}
    We derive the estimate in several steps. 
    \begin{itemize}
        \item 
        Without loss of generality, the barycenter of $S$ is the origin. 
        % Let $I := U \cap T$, which is the union of $k$ different faces of $T$, with $1 \leq k \leq n$.
        %We let $J \subseteq \partial T$ be the union of the remaining $n+1-k$ faces of $T$.
        %So $J$ is within the boundary of the star $|\patch_{\calT}(S)|$ and is disjoint from $S$.
        We fix the subsimplex $S' \subseteq T$ that is complementary to $S$.
        Note that $J$ is the union of exactly those faces of $T$ that contain $S'$.

@@

        \item 
        We introduce a new mapping $\Theta : |\calR| \rightarrow |\calR^c|$.
        Let $K \in \calR$ be an $n$-simplex and $K' \in \calR^{c}$ be constructed from $K$.
        We let $A_{K} : \Delta^{n} \rightarrow K$ and $A_{K'} : \Delta^{n} \rightarrow K'$
        be affine reference transformations
        that agree on the vertices common to $K$ and $K'$.
        We then define 
        \begin{align*}
            \Theta_{|K} := A_{K'} \circ A_{K}^{-1}.
        \end{align*}
        Notice that $A_{K'}$ and $A_{K}$ have the same singular values. Thus
        \begin{align*}
            \| \Jacobian \Theta_{|K}      \| &\leq \| A_{K'} \| \cdot \| A_{K}^{-1} \| \leq \kappa(A_{K}),
            \\
            \| \Jacobian \Theta_{|K}^{-1} \| &\leq \| A_{K'}^{-1} \| \cdot \| A_{K} \| \leq \kappa(A_{K}).
        \end{align*}
        By construction, the mapping preserves volumes:
        \begin{align*}
            |\det( \Jacobian      \Theta_{|K} )|
            = 
            |\det( \Jacobian^{-1} \Theta_{|K} )|
            =
            1
            .
        \end{align*}
        We want to characterize the singular values of this transform. 
        We $h_z \in \bbR^{n}$ be the height vector of $z_{S'}$ inside the simplex $K$.
        We can write
        \begin{align*}
            \Theta_{|K}(x) 
            = 
            x
            - 
            2 \hat z_{S'} \frac{\langle \hat h_z, x \rangle}{\langle \hat h_z, \hat z_{S'} \rangle}
            .
        \end{align*}
        Indeed, we check that the right-hand side equals $x$ whenever $x$ lies in the plane orthgonal to $\hat h_z$,
        and that it equals $-z_{S'}$ when $x = z_{S'}$.
%         and when $x = z_{S'}$, then 
%         \begin{align*}
%             -z_{S'} 
%             =
%             z_{S'}
%             - 
%             2 \hat z_{S'} \frac{\langle \hat h_z, z_{S'} \rangle}{\langle \hat h_z, \hat z_{S'} \rangle}
%             .
%         \end{align*}
        If we orthogonally decompose $z_{S'} = h_z + b_z$, then 
        \begin{align*}
            \Theta_{|K}( h_z ) 
            &= 
            h_z
            - 
            2 z_{S'} \frac{\langle h_z, h_z \rangle}{\langle h_z, z_{S'} \rangle}
            %\\&
            = 
            h_z
            - 
            2 z_{S'}
            %\\&
            = 
            - h_{z}
            - 
            2 b_{z}
            .
        \end{align*}
        Let $\beta$ be the angle between $z_{S'}$ and $h_{z}$. 
        Then $\| h_z \| = \cos(\beta) \|z_{S'}\|$ and $\| b_z \| = \sin(\beta) \|z_{S'}\|$. 
        We study the singular values of the matrix 
        \begin{align*}
            B 
            := 
            \begin{pmatrix}
            -1                 & 0
            \\ 
            -2 \|b_z\|/\|h_z\| & 1
            \end{pmatrix}
            =
            \begin{pmatrix}
            -1             & 0
            \\ 
            -2 \tan(\beta) & 1
            \end{pmatrix}
            .
        \end{align*}
        We compute the eigenvalues of the symmetric matrix $B^{\ast} B$. 
        Its square roots are the desired singular values of $B$:
        \begin{align*}
            \sigma_{\max} = \sqrt{ 2\tan(\beta)^2 + 1 + \tan(\beta) \sqrt{ 4\tan(\beta)^2 + 4 } } = \sqrt{ 1 + \tan^{2}(\beta) } + \tan(\beta)
            ,
            \\
            \sigma_{\min} = \sqrt{ 2\tan(\beta)^2 + 1 - \tan(\beta) \sqrt{ 4\tan(\beta)^2 + 4 } } = \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta)
            .
        \end{align*}
        \begin{align*}
            \lambda
        \end{align*}
        These are also the maximal and minimal singular values of $\Jacobian\Theta_{|K}$. 
        They are monotonely increasing and decreasing, respectively, in $\tan(\beta)$.

%         Let $\beta$ be the angle between $z_{S'}$ and $h_{z}$. 
        By the definition of the tangent, $\tan(\beta) \leq \| b_{z} \| / \| h_{z} \|$, and we know $\|b_z\| \leq \|z_{S'}\|$. 
        Let $F_{z} \subseteq K$ be the face opposite to $z_{S'}$,
        which is also a face of $T$. Then $(n-k) h_{z}$ is the height vector of $F_{z}$ in $T$. 
        Thus 
        \begin{align*}
            \frac{ \| b_{z} \| }{ \| h_{z} \| }
            = 
            \frac{ \| b_{z} \| }{ (n-k)^{-1} \| (n-k) h_{z} \| }
            = 
            \frac{ \| b_{z} \| \vol(F_{z}) }{ (n-k)^{-1} n \vol(T) }
            \leq 
            \frac{ n-k-1 }{n!}
            \cdot 
            \frac{ \diam(T)^{n} }{ \vol(T) }
            :=
            \beta_T
            .
        \end{align*}
        This establishes bounds on the singular values of the transformation in terms of the mesh geometry. 
        We write 
        \begin{align*}
            \beta_T 
            := 
%             \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta) 
%             \geq 
            \sqrt{ 
                1 + (n-k)^{2} \frac{ \diam(T)^{2n} }{ n!^{2}\vol(T)^{2} } 
            } 
            + 
            (n-k) \frac{ \diam(T)^{n} }{ n! \vol(T) }            
            .
        \end{align*}

@@

        We identify the Lipschitz properties of this mapping by computing its Jacobian. 
        We write $\alpha_1, \alpha_2 \geq 0$ for the two angles between $h_1$ and $h_2$, respectively, and the vector $x$.
        In what follows, $y \in \bbR^{n}$.
        \begin{align*}
            \Jacobian F(x) \cdot y
            &=
            \frac{ \langle h_2, h_2 \rangle }{ \|h_2\| \|x\| \cos(\alpha_2) }
            \left( y - \frac{ \langle h_2, y \rangle }{ \langle h_2, x \rangle } x \right)
            \frac{ \langle x, h_1 \rangle }{ \langle h_1, h_1 \rangle }
            +
            x
            \otimes 
            \frac{ \| h_2 \|^{2} }{ \langle h_2, x \rangle }
            \frac{ h_1^{t} y }{ \| h_1 \|^{2} }
            \\&=
            \frac{ \langle \hat h_2, h_2 \rangle }{ \cos(\alpha_2) }
            \left( y - \frac{ \langle \hat h_2, y \rangle }{ \langle \hat h_2, \hat x \rangle } \hat x \right)
            \frac{ \langle \hat x, \hat h_1 \rangle }{ \langle h_1, \hat h_1 \rangle }
            +
            \hat x
            \otimes 
            \frac{ \| h_2 \| }{ \langle \hat h_2, \hat x \rangle }
            \frac{ \hat h_1^{t} y }{ \| h_1 \| }
            \\&=
            \frac{ \| h_2 \| }{ \cos(\alpha_2) }
            \left( y - \frac{ \langle \hat h_2, y \rangle }{ \cos(\alpha_2) } \hat x \right)
            \frac{ \cos(\alpha_1) }{ \| h_1 \| }
            +
            \hat x
            \otimes 
            \frac{ \| h_2 \| }{ \langle \hat h_2, \hat x \rangle }
            \frac{ \hat h_1^{t} y }{ \| h_1 \| }
            \\&=
            \frac{ \| h_2 \| }{ \| h_1 \| }
            \frac{ \cos(\alpha_1) }{ \cos(\alpha_2) }
            \left( 
                y 
                - 
                \frac{ \langle \hat h_2, y \rangle }{ \cos(\alpha_2) } \hat x 
                + 
                \frac{ \langle \hat h_1, y \rangle }{ \cos(\alpha_1) } \hat x 
            \right)
            .
        \end{align*}
        We define $a := \cos(\alpha_1)^{-1} \hat h_1 - \cos(\alpha_2)^{-1} \hat h_2$. 
        Notice that $a \perp \hat x$ by definition. The matrix $\Id + \hat x \otimes a^{t}$ is the identity over the orthogonal complement of the linear span of $a$ and $x$. Over that two-dimensional subspace, it can be represented by a triangular matrix.
        Similar as above, its largest and the smallest singular values $\sigma_{max} \geq 1 \geq \sigma_{\min}$ are
        %~\cite{blinn1996consider} 
        \begin{align*}
            \sigma_{\max} = \sqrt{ 1 + \frac{ \|a\|^{2} }{ 4 } } + \frac{ \|a\| }{2},
            \qquad 
            \sigma_{\min} = \sqrt{ 1 + \frac{ \|a\|^{2} }{ 4 } } - \frac{ \|a\| }{2}.     
        \end{align*}
        Notice that these are strictly monotonely increasing or decreasing, respectively, in $\|a\|$.
        We notice the obvious upper bound 
        \begin{align*}
            \|a\| 
            \leq 
            2 \max_{1 \leq i \leq 2} 
            \left( 
                \frac{ \hat h_i }{ \langle \hat x, \hat h_i \rangle } 
            \right)
            .
        \end{align*}

@@

        The diameters of the simplices in $\calK$ are already bounded by the maximum diameter in $\patch_{\calT}(S)$.
        The height $h_{K}$ is obtained from a height vector of a vertex in $\patch_{\calT}(S)$ by scaling with the factor $(k+1)^{-1}$.
        Every height of the origin in some simplex of $\calR'$ is obtained from a height vector of the origin in some simplex in $\calS'$
        by scaling with $(k+1)^{-1}$. Each of the latter height vectors has length bounded from below by the height of some vertex in $S$ within a simplex of $\calS$, multiplied by $\beta_T$. The height of any vertex of $S$ in some simplex of $\calS$ is the same as the height $h_T$ of that vertex in the original simplex $T$. 
        In oder to estimate $\|x_{G}\|$, recall that $x = \lambda (-z_{S'}) + (1-\lambda) z_{x}$ for some $\lambda \in [0,1]$ and $z_{x} \in G$
        being in the face of $G$ that is opposite to $z_{x}$. The preimage of $x$ under $\Theta_{{G}}$ is $\lambda z_{S'} + (1-\lambda) z_{x}$. 
        It follows that $\|x_{G}\| \leq \diam(T)$.

        We know that whenever $T$ is a simplex with some face $F$ and $h_F$ is corresponding height vector, then 
        \begin{align*}
            \frac{\diam(T)}{ \|h_{F}\|}
            \leq  
            \frac{ \vol(F) \diam(T) }{ n \vol(T) }
            \leq 
            \frac{ \diam(T)^{n} }{ n! \vol(T) }
            .
        \end{align*}
        We conclude that 
        \begin{gather*}
            \frac{ \diam(K) }{ \| h_{{G}} \| }
            \leq 
            (k+1)\beta_T
            \frac{ \diam(K) }{ \| h_{T} \| }
            \leq 
            (k+1)\beta_T
            \theta(\calT)
            \frac{ \diam(T) }{ \| h_{T} \| }
            \leq 
            (k+1)\beta_T
            \theta(\calT)
            \frac{ \diam(T)^{n} }{ n! \vol(T) }
            ,
            \\
            \frac{ \| x_{G} \| }{ \| h_{K} \| }
            \leq 
            (k+1)
            \sup_{ K \in \patch_{\calT}(S) }
            \frac{ \diam(T) }{ \| h_{K} \| }
            \leq 
            (k+1)
            \theta(\calT)
            \sup_{ K \in \patch_{\calT}(S) }
            \frac{ \diam(K) }{ \| h_{K} \| }
            \leq 
            (k+1)
            \theta(\calT)
            \sup_{ K \in \patch_{\calT}(S) }
            \frac{ \diam(K)^{n} }{ n! \vol(K) }
            .
        \end{gather*}
        \begin{gather*}
            \frac{ \diam(K) }{ \| h_{{K}} \| }
            \leq 
            (k+1)
            \sup_{ K \in \patch_{\calT}(S) }
            \frac{ \diam(K)^{n} }{ n! \vol(K) }
            ,
            \qquad 
            \frac{ \| x_{G} \| }{ \| h_{G} \| }
            \leq 
            (k+1)\beta_T
            \frac{ \diam(T)^{n} }{ n! \vol(T) }
            .
        \end{gather*}

@@

        The vector $a$ is the difference between $\hat h_{G}$ and $\hat h_{K}$ after both are rescaled such that they lie on the affine hyperplane with normal vector $\hat x$. That difference has length at most $\tan(\alpha_{G}) + \tan(\alpha_{K})$.
        We calculate 
        \begin{align*}
            \|a\|
            = 
            \left\|
            \frac{\|x_{G}\|}{\|h_{G}\|} \hat h_{{G}} 
            - 
            \frac{\|x_{K}\|}{\|h_{K}\|} \hat h_K
            \right\|
            &
            \leq 
            \frac{\|x_{G}\|}{\|h_{G}\|} 
            + 
            \frac{\|x_{K}\|}{\|h_{K}\|} 
            \\&
            \leq 
            (k+1) 
            \left( 1 + \beta_T \right)
            \sup_{ K \in \patch_{\calT}(S) }
            \frac{ \diam(K)^{n} }{ n! \vol(K) }
            .
        \end{align*}

@@

        We recall the upper bound $\sqrt{1+\theta^2} \leq 1 + \theta$ for any $\theta \geq 0$. 
        This leads to  
        \begin{align*}
            \| \Jacobian \Xi \|
            &\leq 
            \Cvier{n}{k}(\calT)
            :=
            (k+1) 
            \mu(\calT)
            \beta_T
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + \beta_T \right)
                \mu(\calT)
            \Big)            
            \\
            \| \Jacobian \Xi^{-1} \|
            &\leq  
            \Cfive{n}{k}(\calT)
            :=
            (k+1) 
            \mu(\calT)
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + \beta_T \right)
                \mu(\calT)
            \Big)            
        \end{align*}
        We get the determinant estimate 
        \begin{align*}
            |\det(\Xi)|
            \leq 
            \Cdetvier{n}{k}(\calT)
            :=
            (k+1)^{n}
            \theta(\calT)^{n}
            \beta_T^{n}
            \mu(\calT)^{n}
            ,
            \qquad 
            |\det(\Xi^{-1})|
            \leq 
            \Cdetfive{n}{k}(\calT)
            :=
            (k+1)^{n}
            \theta(\calT)^{n}
            \mu(\calT)^{n}
            .
        \end{align*}
        This completes the desired estimates.
